Weather Data Integration and Assimilation System\\

### 2020-02-11 lit/sec_1-fews.tex

\begin{figure}[htp]
    \centering
    \includegraphics[width=1.0\textwidth]{lit/fews/Architecture-of-Delft-FEWS-showing-the-data-base-the-data-access-layers-and-examples-of_W640.png}
    \caption[Architecture of \acrshort{fews}]{Architecture of \acrshort{fews} \cite{Werner2013TheSystem}.}
    \label{fi:fews_data_layer}
\end{figure}
Most of the library of functions provided work equally on scalar and gridded time series. As with all other modules this communicates with the database solely through the data access layer as shown in \db{Section} \ref{fi:fews_data_layer} \cite{Werner2013TheSystem}. As it shown in the figure, it is clearly showing that, the system is depends on single database and via the data access layer all the requests are coming to the database though a connection. It is possible to setup and connect to a enterprise level database with clustering and shading as a paid solution, to serve many requests as possible. But the design it self inherently suffer with the database bottleneck which effected on running a large set of timeseries or models.
\dbc{When referring to a section you need to use the word "Section" as highlighted above. Fix all other places.}

>>>>

\begin{figure}[htp]
    \centering
    \includegraphics[width=1.0\textwidth]{lit/fews/Architecture-of-Delft-FEWS-showing-the-data-base-the-data-access-layers-and-examples-of_W640.png}
    \caption[Architecture of \acrshort{fews}]{Architecture of \acrshort{fews} \cite{Werner2013TheSystem}.}
    \label{fi:fews_data_layer}
\end{figure}
Most of the library of functions provided work equally on scalar and gridded time series. As with all other modules this communicates with the database solely through the data access layer as shown in \hl{Section} \ref{fi:fews_data_layer} \cite{Werner2013TheSystem}. As it shown in the figure, it is clearly showing that, the system is depends on single database and via the data access layer all the requests are coming to the database though a connection. It is possible to setup and connect to a enterprise level database with clustering and shading as a paid solution, to serve many requests as possible. But the design it self inherently suffer with the database bottleneck which effected on running a large set of timeseries or models.
\dbc{When referring to a section you need to use the word "Section" as highlighted above. Fix all other places.}

### 2020-02-11 method/sec_0_introduction.tex

\subsection{Modules of Weather Data System}
\label{subse:modules_weather_data_integration_sys}

Figure \ref{fi:wdia_components} shows the basic functions of a Weather data integration and assimilation system which consists of the following modules:
\begin{figure}[htbp]
\centerline{\includegraphics[width=0.8\textwidth]{method/misc/weather_data_system_components.jpg}}
\caption{Components of a weather data system.}
\label{fi:wdia_components}
\end{figure}

@@

\paragraph{Timeseries}-- A timeseries is simply a series of data points ordered in time. In weather domain, it interest in timeseries in perspective of observations to forecasting. Each timeseries, the data points can be form in different formats as well. For example, Scalar (0D), Vector (1D), Grid (2D), and Polygon (2D).

WDIAS is focus on handling Scalar, Vector and Grid data timeseries via the thesis. But the system is capable of extending to handle polygon timeseries as well.

\subsubsection{Base of WDIAS architecture}
\dbc{What's Base of WDIAS architecture?}
The base of WDIAS architecture origins based on attributes of a timeseries. There are many attributes to differentiate timeseries one from another. But following can be consider as key attributes to uniquely identify a timeseries from another for the \acrshort{wdias}. In detail description available at the section \ref{se:db_struct}.
\begin{itemize}
    \item Module -- module that generate the timeseries
    \item Value type -- Scalar, Vector, Grid, Polygon
    \item Location 
    \item Parameter -- measuring parameter of timeseries
    \item Timeseries type
    \begin{itemize}
        \item sources -- External, Simulated
        \item category -- Historical, forecast
    \end{itemize}
    \item Time step
\end{itemize}
\dbc{Use bullets only when essential. Other places use paragraphs.}

>>>>

\subsection{Modules of a Weather Data System}
\label{subse:modules_weather_data_integration_sys}

Figure \ref{fi:wdia_components} shows the basic functions of a Weather data integration and assimilation system which consists of the following modules:
\begin{figure}[htbp]
\centerline{\includegraphics[width=0.8\textwidth]{method/misc/weather_data_system_components.jpg}}
\caption{Modules of a weather data system.}
\label{fi:wdia_components}
\end{figure}

@@

\paragraph{Timeseries}-- A timeseries is simply a series of data points ordered in time. In weather domain, it interest in timeseries in perspective of observations to forecasting. Each timeseries, the data points can be form in different formats as well. For example, scalar (0D), vector (1D), grid (2D), and polygon (2D).
dbc{Why are you referring to timeseries here? It's not one of the 3 key components in Fig. 3.1.}

In the proof of concept design described below, WDIAS focuses on handling scalar, vector and grid timeseries data. However, the system could be extended to handle polygon-based timeseries as well.

\subsubsection{Base of WDIAS architecture}
\dbc{What's Base of WDIAS architecture?}
\hl{The base of WDIAS architecture origins based on attributes of a timeseries.} %There are many attributes to differentiate timeseries one from another. But 
Among many attribute to differentiate a timeseries from another, the following can be consider as the key attributes in uniquely identify a timeseries from another: % for the \acrshort{wdias}.
\dbc{Highlighted sentence doesn't make sense to me. All yellow highlights mean they are not clear.}

\begin{itemize}
    \item Module -- module that generates the timeseries
    \item Value type -- Scalar, vector, grid, or polygon
    \item Location of readings
    \item Parameter -- measuared parameter
    \item hl{Timeseries type}
    \begin{itemize}
        \item Sources -- External or simulated
        \item Category -- Historical or forecast
    \end{itemize}
    \item Time step
\end{itemize}
\dbc{Use bullets only when essential. Other places use paragraphs.}

A detail description of the above attributes are presented in Section \ref{se:db_struct}.

### 2020-02-11 method/sec_1-architectural_decisions.tex

\section{Architectural Desicions}
\label{se:architectural_decisions}

Initial design of \acrshort{wdias} used the \acrfull{soa}, and try to use the \acrfull{esb} which can provide capability to integrate different modules, since it is acting as a common layer for all of modules. By default, \acrshort{esb} provides pub sub capabilities.
Since \acrshort{esb} mediator can be used to integrate the modules which are described in \ref{subse:modules_weather_data_integration_sys}. \acrshort{esb} also support different transportation protocols such as HTTP and Web-Socket etc.

But \acrshort{esb} is not a better solution for data streaming or bulk data processing. Also \acrshort{esb} suffer from single point of failure, since all the messages are going though a common bus.

In the second phase of \acrshort{wdias}, the proposed architecture changed to use with Actor model using AKKA framework \cite{HewittWhyModel} after consider into the disadvantage of using \acrshort{esb}. Using the Actors we tried to implement the microservice architecture. In that case, an Actor or An Actor with slave Actors map into each microservice in the \acrshort{wdias}.

The typical \acrshort{soa} model, for example, usually has more dependent \acrshort{esb}s, with microservices using faster messaging mechanisms. \acrshort{soa} also focuses on imperative programming, whereas microservices architecture focuses on a responsive-actor programming style. Moreover, \acrshort{soa} models tend to have an outsized relational database, while microservices frequently use NoSQL or micro-SQL databases (which can be connected to conventional databases). But the real difference has to do with the architecture methods used to arrive at an integrated set of services in the first place.

As mentioned above, \acrshort{esb} support to implement \acrshort{soa}. But \acrshort{soa} has limitations such as single point of failure, slower communication (can’t use for transfer data) etc. So, the microservice architecture get evolved and demanding for the moment. When compare to \acrshort{soa}, microservice architecture has some advantages;
\begin{itemize}
    \item Follow the Single Responsibility Principle
    \item Resilient/Flexible – failure in one service does not impact other services. If you have monolithic or bulky service errors in one service/module it can impact other modules/functionality.
    \item High scalability – demanding services can be deployed in multiple servers to enhance performance and keep away from other services so that they don’t impact other services. Will be difficult to achieve the same with single, large monolithic service.
    \item Easy to enhance – less dependency and easy to change and test
    \item Low impact on other services – being an independent service, this has less chance to impact other services
    \item Easy to understand since they represent the small piece of functionality
    \item Ease of deployment
\end{itemize}

Even with using AKKA framework, it has some of the disadvantage as described in AKKA documentation \cite{Akka.ioWhenCluster}. Even Actor model capable of implementing a microservice architecture, it has few disadvantage. In order to overcome these issues, we moved to the \acrfull{k8s} \cite{LinuxFoundationProduction-GradeKubernetes}.;
\begin{itemize}
    \item Unable to choose technology. But \acrshort{k8s} allows you to choose technology that is best suited for a particular functionality
    \item As mentioned in the AKKA documentation \cite{Akka.ioWhenCluster}, it's better for developing one microservice within the whole \acrshort{wdias} design.
\end{itemize}

>>>>

\section{Architectural Decisions}
\label{se:architectural_decisions}

Initial design of \acrshort{wdias} used the \acrfull{soa}. Moreover, it tried to use an \acrfull{esb} to integrate different modules, as it is acting as a common layer for all of modules. By default, \acrshort{esb} provides publisher/subscribe (pub/sub) capabilities.
\hl{Because} \acrshort{esb} \hl{mediator can be used to integrate the modules which are described in Section} \ref{subse:modules_weather_data_integration_sys}. \acrshort{esb} also support different transportation protocols such as HTTP and Web sockets. However, it turned out that \acrshort{esb} is not a suitable for data streaming or bulk data processing. Also, \acrshort{esb} suffer from single point of failure, as all the messages are going though a common bus.

In the second design round of \acrshort{wdias}, we attempted to use the actor model using AKKA framework \cite{HewittWhyModel}. %However, after consider into the disadvantage of using \acrshort{esb}. 
Using the actors we tried to implement the microservice architecture. In that case, an actor or an actor with slave actors were map to each microservice in \acrshort{wdias}.

/hl{The typical \acrshort{soa} model, for example, usually has more dependent \acrshort{esb}s, with microservices using faster messaging mechanisms. \acrshort{soa} also focuses on imperative programming, whereas microservices architecture focuses on a responsive-actor programming style. Moreover, \acrshort{soa} models tend to have an outsized relational database, while microservices frequently use NoSQL or micro-SQL databases (which can be connected to conventional databases). But the real difference has to do with the architecture methods used to arrive at an integrated set of services in the first place.}
\dbc{Multiple design rounds and decisions are mixed in the para. First finish discussing ESB. Then finish actor model in another para. Then focus on mircoservice details.}

%As mentioned above, \acrshort{esb} support to implement \acrshort{soa}. But \acrshort{soa} has limitations such as single point of failure, slower communication (can’t use for transfer data) etc. So, the microservice architecture get evolved and demanding for the moment. 
When compare to \acrshort{soa}, the microservice architecture has several advantages such as the following:
\begin{itemize}
    \item Follow the Single Responsibility principle
    \item Resilient/flexible as failure in one service does not impact other services. If you have monolithic or bulky service errors in one service/module it can impact other modules/functionality.
    \item High scalable as demanding services can be deployed in multiple servers to enhance performance and keep away from other services so that they do not impact other services. Will be difficult to achieve the same with single, large monolithic service.
    \item Easy to enhance as dependencies are less. Is also easier to change and test.
    \item Low impact on other services as each service is independent. % service., this has less chance to impact other services
    \item Easy to understand since they represent a small piece of functionality
    \item Ease of deployment
\end{itemize}

Even with using the AKKA framework, it has some of the disadvantage as described in AKKA documentation \cite{Akka.ioWhenCluster}. Even Actor model capable of implementing a microservice architecture, it has few disadvantage. To overcome these issues, we moved to the \acrfull{k8s} \cite{LinuxFoundationProduction-GradeKubernetes}.
\dbc{Kubernetes is a product. What you need to refer is to the concept of container-orchestration. Fix.}

\begin{itemize}
    \item \hl{Unable to choose technology. But} \acrshort{k8s} \hl{allows you to choose technology that is best suited for a particular functionality}
    \item \hl{As mentioned in the AKKA documentation} \cite{Akka.ioWhenCluster}, \hl{it is better for developing one microservice within the whole} \acrshort{wdias} design.
\end{itemize}

\dbc{What are these 2 bullets?}
\dbc{Don't use shortened words like it's, isn't, don't,... in writing}

### 2020-02-11 method/sec_4-db_struct.tex

The database structure of WDIAS mainly depends on the key attributes of the timeseries. A brief summary of possible value for timeseries can be describe as below;

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Timeseries Key Attributes}
\label{subse:timeseries_key_attributes}
\subsubsection{Module ID}
String field which describe the source of the data generated. e.g., hec-hms, flow2d, weather-station etc

\subsubsection{Value Type}
Scalar, Vector, Grid are the Value Types which are interested in the \acrshort{wdias}.

\subsubsection{Location}
Location of the timeseries. All locations have unique String identifier called locationId. Further, there are two type of location types that is interested in the WDIAS.
\begin{itemize}
  \item \emph{Point locations} - Contains a name which is human readable. lat and lon of the location on earth surface.
  \begin{lstlisting}[language=Python]
    {
        "locationId": "wdias-hanwella",
        "name": "Hanwella",
        "lat": 6.909722222,
        "lon": 80.08166667
    }
  \end{lstlisting}
  \item \emph{Regular Grid locations} - Contains a name which is human readable. The grid presents by dividing into equal size cells. Thus, it need the number of rows and columns. 
  And the location of the first cell and the width and height of it.
  \begin{lstlisting}[language=Python]
      {
        "locationId": "wdias_kelani_basin",
        "description": "Kelani Basin",
        "rows": 120,
        "columns": 139,
        "geoDatum": "Kandawala",
        "gridFirstCell": {
            "firstCellCenter": {
                "x": 397074.0,
                "y": 504875.0
            },
            "xCellSize": 250.0,
            "yCellSize": 250.0
        }
    }
  \end{lstlisting}
  \item \emph{Irregular Grid locations} - (endpoints are available with in the WDIAS system, but skipped since it is out of the interest of the scope.)
\end{itemize}

\subsubsection{Parameter}
Parameter describe the variable measuring against a location. All parameters ahve a unique String identifier called as parameterId. A parameter can be use in multiple locations.
While defining a parameter, there are three required fields which need to be provided such as variable, unit and parametertype.
\begin{itemize}  
  \item \emph{Variable} - Nature of the variable measuring. Example Precipitation, Temperature and WaterLevel etc
  \item \emph{Unit} - metric unit of measuring
  \item \emph{Parameter Type} - Should be one of 'Instantaneous', 'Accumulative', 'Mean'
  \begin{lstlisting}[language=Python]
    {
      "parameterId": "O.Precipitation",
      "variable": "Precipitation",
      "unit": "mm",
      "parameterType": "Instantaneous"
    }
  \end{lstlisting}
\end{itemize}

\subsubsection{Timeseries Type}
Timeseries Type can be combination of Source followed by Category.
\begin{itemize}
  \item Sources - External, Simulated
  \item Category - Historical, Forecast
\end{itemize}

\subsubsection{Time Step}
All Time Steps has an uniqie String identifier called as timeStepId. Unit should be one of 'Second', 'Minute', 'Hour', 'Day', 'Week', 'Month', 'Year', 'NonEquidistant'. One of multiplier or divider can be use to define the interval between each measurement.
\begin{lstlisting}[language=Python]
{
    "timeStepId": "each_min",
    "unit": "Minute",
    "multiplier": 1,
    "divider": 0
}
\end{lstlisting}

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\acrshort{rdbms}}
\label{subse:mysql}

There are lots of \acrshort{rdbms} available and for \acrshort{wdias} is using SQL as the \acrfull{rdbms}.
SQL database has fewer version of open source and it is also possible to gain better performance with commercial versions. 
Since WDIAS scope is to using open source tools, it is using \acrshort{mysql} as persistent database for storing timeseries metadata described above.
One instance of \acrshort{mysql} is using only by adapter-metadata service.
Other than that, another instance of \acrshort{mysql} is using for adapter-extension which will be describe in more detail \ref{se:data_preprocess}.
\acrshort{mysql} is used only for these services, those will only keep the consistency of the schema for timeseries metadata and extension metadata.
\ref{subse:redis} instance is using for caching the data from \acrshort{mysql}, since the \acrshort{mysql} database instances, itself not going to get much load.
Other than MySQL, there are many open source databases can be used to replace the same functionality.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Timeseries Database}
\label{subse:influxdb}
Among the other timeseries open source databases available, \acrshort{influxdb} \cite{influxdbInfluxDBDocumentation} has higher number of usage and support. Database is written in Golang programming language, and has higher performance, and has following supports;
\begin{itemize}
  \item Open timeseries DB with MIT license
  \item Support SQL-Like query language
  \item Clustering available with a commercial version
\end{itemize}
Other than \acrshort{influxdb}, another Database is Elastic Search which is mainly using for indexing for searching.
With changing the configurations of the \acrshort{influxdb}, it is possible to adapt as required. As mention in \cite{influxdbInfluxDBDocumentation}, 
\acrshort{influxdb} can setup on a computer which will give the required performance. Going further, users can use commercial version of \acrshort{influxdb} in order to gain more performance.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\acrfull{netCDF}} 
\label{subse:netcdf}
\acrshort{netCDF} \cite{unidataUnidataNetCDF} is a self-describing, machine-independent data formats that support creation, access, and sharing of array-oriented scientific data.
And it also has the support for parallel file access.
NetCDF is widely using in sciencetific domain due to it's capability of storing data in multi dementions easily and flexibility with storing many of the sciencetific data.
Python wrapper for NetCDF-C library using in the WDIAS system, and use by adapter-grid service.

@@

\acrshort{wdias} is using \acrshort{mongodb} as the Document-oriented databse, since it has the Geo searching capabilities.
\acrshort{mongodb} \cite{mongodbMongoDBManual} is a general purpose, document-based, distributed database. And it supports;
\begin{itemize}
  \item Supports query operations on geo-spatial data \cite{mongodbMongoDBManual}
  \item Clustering and Sharding available for scalability and reliability
\end{itemize}
\acrshort{mongodb} geo-spatial features are using for support Geo queries in the WDIAS. Other than that, external timeseries metadata queries are served by the indexed timeseries metadata data.
If metadata not present in the adapter-query, if fetch data from the adapter-metadata, then indexed and cache in the \acrshort{mongodb}. Vise a versa, if new timeseries create in the adapter-metadata, it trigger an event to adapter-query to index the new timeseries.
In Section \ref{subse:scale_cube}, the Y-axis scaling concept is using here by splitting into the multiple services based on the functionality and scale the system.

>>>>

The database structure of WDIAS mainly depends on the key attributes of the timeseries. \hl{A brief summary of possible value for timeseries can be describe as below};
\dbc{Clerly mention why a hierarchical design is needed.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Key Attributes of Timeseries}
\label{subse:timeseries_key_attributes}
\paragraph{Module ID}-- String field which describe the source of the data generated. e.g., hec-hms, flow2d, weather-station etc

\paragraph{Value Type}-- Scalar, Vector, Grid are the Value Types which are interested in the \acrshort{wdias}.

\paragraph{Location}-- Location of the timeseries. All locations have unique String identifier called locationId. Further, there are two type of location types that is interested in the WDIAS, namely:
\begin{itemize}
  \item \emph{Point locations}-- Contains a name which is human readable. lat and lon of the location on earth surface.
  \begin{lstlisting}[language=Python]
    {
        "locationId": "wdias-hanwella",
        "name": "Hanwella",
        "lat": 6.909722222,
        "lon": 80.08166667
    }
  \end{lstlisting}
  \dbc{Remove additional line.}

  \item \emph{Regular Grid locations} -- Contains a name which is human readable. The grid presents by dividing into equal size cells. Thus, it need the number of rows and columns. 
  And the location of the first cell and the width and height of it.
  \begin{lstlisting}[language=Python]
      {
        "locationId": "wdias_kelani_basin",
        "description": "Kelani Basin",
        "rows": 120,
        "columns": 139,
        "geoDatum": "Kandawala",
        "gridFirstCell": {
            "firstCellCenter": {
                "x": 397074.0,
                "y": 504875.0
            },
            "xCellSize": 250.0,
            "yCellSize": 250.0
        }
    }
  \end{lstlisting}
  \item \emph{Irregular Grid locations} -- (endpoints are available with in the WDIAS system, but skipped since it is out of the interest of the scope.)
\end{itemize}

\paragraph{Parameter}
Parameter describe the variable measuring against a location. All parameters ahve a unique String identifier called as \textit{parameterId}. A parameter can be use in multiple locations.
While defining a parameter, there are three required fields which need to be provided such as variable, unit and \textit{parametertype}.
\begin{itemize}  
  \item \emph{Variable} -- Nature of the variable measuring. Example Precipitation, Temperature and WaterLevel etc
  \item \emph{Unit} -- metric unit of measuring
  \item \emph{Parameter Type} -- Should be one of Instantaneous, Accumulative, or Mean
  \begin{lstlisting}[language=Python]
    {
      "parameterId": "O.Precipitation",
      "variable": "Precipitation",
      "unit": "mm",
      "parameterType": "Instantaneous"
    }
  \end{lstlisting}
\end{itemize}

\subsubsection{Timeseries Type}
Timeseries Type can be combination of Source followed by category:
\begin{itemize}
  \item Sources - External or simulated
  \item Category - Historical or forecast
\end{itemize}
\dbc{Explain each with sufficient detail.}

\paragraph{Time Step}-- All Time Steps has an uniqie String identifier called as timeStepId. Unit should be one of Second, Minute, Hour, Day, Week, Month, Year, or NonEquidistant. One of multiplier or divider can be use to define the interval between each measurement.
\begin{lstlisting}[language=Python]
{
    "timeStepId": "each_min",
    "unit": "Minute",
    "multiplier": 1,
    "divider": 0
}
\end{lstlisting}

@@

\dbc{Fig. 3.6 needs a more meaningful title.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\acrshort{rdbms}}
\label{subse:mysql}
\dbc{Needs a more meaningful Sec. title}

%There are lots of \acrshort{rdbms} available and for 
\acrshort{wdias} is using SQL as the \acrfull{rdbms}. 
%SQL database has fewer version of open source and it is also possible to gain better performance with commercial versions. 
Since WDIAS scope is to using open source tools, it is using \acrshort{mysql} as persistent database for storing timeseries metadata described above.
One instance of \acrshort{mysql} is using only by adapter-metadata service.
Other than that, another instance of \acrshort{mysql} is using for adapter-extension which will be describe in more detail \ref{se:data_preprocess}.
\acrshort{mysql} is used only for these services, those will only keep the consistency of the schema for timeseries metadata and extension metadata.
\ref{subse:redis} instance is using for caching the data from \acrshort{mysql}, since the \acrshort{mysql} database instances, itself not going to get much load.
Other than MySQL, there are many open source databases can be used to replace the same functionality.
\dbc{First start with a brief discussion on why a RDBMS is needed}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Timeseries Database}
\label{subse:influxdb}
Among several timeseries open source databases, \acrshort{influxdb} \cite{influxdbInfluxDBDocumentation} \hl{has higher number of usage and support. Database is written in Golang programming language, and} has higher performance, and has following supports;
\begin{itemize}
  \item Open timeseries DB with MIT license
  \item Support SQL-Like query language
  \item Clustering available with a commercial version
\end{itemize}
Other than \acrshort{influxdb}, another database is Elastic Search which is mainly using for indexing for searching.
With changing the configurations of the \acrshort{influxdb}, it is possible to adapt as required. As mention in \cite{influxdbInfluxDBDocumentation}, 
\acrshort{influxdb} can setup on a computer which will give the required performance. Going further, users can use commercial version of \acrshort{influxdb} in order to gain more performance.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\acrfull{netCDF}} 
\label{subse:netcdf}
\acrshort{netCDF} \cite{unidataUnidataNetCDF} is a self-describing, machine-independent data formats that support creation, access, and sharing of array-oriented scientific data.
And it also has the support for parallel file access.
NetCDF is widely using in sciencetific domain due to it's capability of storing data in multi dementions easily and flexibility with storing many of the sciencetific data. \hl{
Python wrapper for NetCDF-C library using in the WDIAS system, and use by adapter-grid service.}

@@

\acrshort{wdias} is using \acrshort{mongodb} as the document-oriented databse, as it has the Geo searching capabilities.
\acrshort{mongodb} \cite{mongodbMongoDBManual} is a general purpose, document-based, distributed database. And it supports:
\dbc{Need for a mix of relational, time series, document stores, and in-memory need to be first mentioned and justified in high-level design of EDIAS above.}
\begin{itemize}
  \item Supports query operations on geo-spatial data \cite{mongodbMongoDBManual}
  \item Clustering and Sharding available for scalability and reliability
\end{itemize}
\acrshort{mongodb} geo-spatial features are using for support Geo queries in the WDIAS. Other than that, external timeseries metadata queries are served by the indexed timeseries metadata data.
If metadata not present in the adapter-query, if fetch data from the adapter-metadata, then indexed and cache in the \acrshort{mongodb}. Vise a versa, if new timeseries create in the adapter-metadata, it trigger an event to adapter-query to index the new timeseries.
In Section \ref{subse:scale_cube}, the Y-axis scaling concept is using here by splitting into the multiple services based on the functionality and scale the system.

### 2020-02-11 method/sec_6-query.tex
Fixed: sub section labels
Misc fixes on references

As described in \db{Section} \ref{sub:mongodb}, adapter-query is primarily using for support Geo-spatial queries. Other than that, as the primary service for searching for timeseries, it also supports for retrieval of timeseries metadata for external queries as well.

\subsection{Timeseries Metadata}
Get timeseries metadata by timeseriesID.
\begin{lstlisting}
    GET <HOST_NAME>/metadata/timeseries/<TIMESERIES_ID>
\end{lstlisting}
\dbc{Use monospace formatting.}

Search for timeseries metadata with key attributes of the timeseries.
\begin{lstlisting}
    GET <HOST_NAME>/metadata/timeseries?<QUERY_STRING>
\end{lstlisting}
QUERY\_STRING can be consist of zero to multiple parameters. Multiple query parameters are separated by the ampersand, "\&".
\begin{itemize}
    \item moduleId(Optional), e.g., HEC-HMS
    \item valueType(Optional), e.g., Scalar
    \item parameterId(Optional), e.g., O.Precipitation
    \item locationId(Optional), e.g., wdias\_hanwella
    \item timeseriesType(Optional), e.g., ExternalHistorical
    \item timeStepId(Optional), e.g., each\_15\_min
\end{itemize}
For the moment, query are only support by strict matching with above parameter values.

@@

\subsubsection{Query Locations within Area}
Using following API endpoint, it is possible to query the locations which are exists within the area provided by the geoJson data in the JSON payload.
\begin{lstlisting}
    POST <HOST_NAME>/query/location
    JSON Body:
    {
        "geoJson": {
            "type": "Polygon",
            "coordinates": [
                [
                    [
                        <longitude>,
                        <latitude>
                    ],
                    ...
                ]
            ]
        }
    }
\end{lstlisting}

\subsubsection{Query parameters in Locations}
Using the following API endpoint, it is possible to query the parameters stored against given locations. Location IDs should be provide as list of JSON payload.
\begin{lstlisting}
    POST <HOST_NAME>/query/parameter
    JSON Body:
    ["<locationID>", ...]
\end{lstlisting}

\subsubsection{Query Timeseries by Location}
Using the following API endpoint, it is possible to query the timeseries metadata for a given location.
\begin{lstlisting}
    POST <HOST_NAME>/query/timeseries
    JSON Body:
    {
        "location": "<LOCATION_ID>"
    }
\end{lstlisting}

@@

\subsubsection{Query Timeseries in Area}
Using the following API endpoint, it is possible to query the timeseries metadata within a given area. This operation uses geoWithin \acrshort{mongodb} operator for the simplicity.
\begin{lstlisting}
    POST <HOST_NAME>/query/timeseries
    JSON Body:
    {
        "geoJson": {
            "type": "Polygon",
            "coordinates": [
                [
                    [
                        <longitude>,
                        <latitude>
                    ],
                    ...
                ]
            ]
        }
    }
\end{lstlisting}

\subsubsection{Query Timeseries in Area by Parameter}
Using the following API endpoint, it is possible to query the timeseries metadata within a given area for given parameter. This is an extension to the above endpoint.
\begin{lstlisting}
    POST <HOST_NAME>/query/timeseries
    JSON Body:
    {
        "geoJson": {
            "type": "Polygon",
            "coordinates": [
                [
                    [
                        <longitude>,
                        <latitude>
                    ],
                    ...
                ]
            ]
        },
        "parameter": "<PARAMETER_ID>"
    }
\end{lstlisting}

\subsubsection{Query All Timeseries}
Send a request with empty JSON body will respond with all the timeseries.
\begin{lstlisting}
    POST <HOST_NAME>/query/timeseries
    JSON Body:
    {}
\end{lstlisting}

>>>>

As described in Section \ref{subse:mongodb}, adapter-query is primarily using for support Geo-spatial queries. Other than that, as the primary service for searching for timeseries, it also supports for retrieval of timeseries metadata for external queries as well.

\subsection{Timeseries Metadata}
Get timeseries metadata by timeseriesID.
\begin{lstlisting}
    GET <HOST_NAME>/metadata/timeseries/<TIMESERIES_ID>
\end{lstlisting}

Search for timeseries metadata with key attributes of the timeseries.
\begin{lstlisting}
    GET <HOST_NAME>/metadata/timeseries?<QUERY_STRING}
\end{lstlisting}

QUERY\_STRING can be consist of zero to multiple parameters. Multiple query parameters are separated by the ampersand, "\&".
\begin{itemize}
    \item moduleId(Optional), e.g., HEC-HMS
    \item valueType(Optional), e.g., Scalar
    \item parameterId(Optional), e.g., O.Precipitation
    \item locationId(Optional), e.g., wdias\_hanwella
    \item timeseriesType(Optional), e.g., ExternalHistorical
    \item timeStepId(Optional), e.g., each\_15\_min
\end{itemize}
For the moment, query are only support by strict matching with above parameter values.

@@

\paragraph{Query Locations within Area}-- Using following API endpoint, it is possible to query the locations which are exists within the area provided by the geoJson data in the JSON payload.
\begin{lstlisting}
    POST <HOST_NAME>/query/location
    JSON Body:
    {
        "geoJson": {
            "type": "Polygon",
            "coordinates": [
                [
                    [
                        <longitude>,
                        <latitude>
                    ],
                    ...
                ]
            ]
        }
    }
\end{lstlisting}

paragraph{Query parameters in Locations}-- Using the following API endpoint, it is possible to query the parameters stored against given locations. Location IDs should be provide as list of JSON payload.
\begin{lstlisting}
    POST <HOST_NAME>/query/parameter
    JSON Body:
    ["<locationID>", ...]
\end{lstlisting}

\paragraph{Query Timeseries by Location}-- Using the following API endpoint, it is possible to query the timeseries metadata for a given location.
\begin{lstlisting}
    POST <HOST_NAME>/query/timeseries
    JSON Body:
    {
        "location": "<LOCATION_ID>"
    }
\end{lstlisting}

@@

\paragraph{Query Timeseries in Area} -- Using the following API endpoint, it is possible to query the timeseries metadata within a given area. This operation uses geoWithin \acrshort{mongodb} operator for the simplicity.
\begin{lstlisting}
    POST <HOST_NAME>/query/timeseries
    JSON Body:
    {
        "geoJson": {
            "type": "Polygon",
            "coordinates": [
                [
                    [
                        <longitude>,
                        <latitude>
                    ],
                    ...
                ]
            ]
        }
    }
\end{lstlisting}

\paragraph{Query Timeseries in Area by Parameter}-- Using the following API endpoint, it is possible to query the timeseries metadata within a given area for given parameter. This is an extension to the above endpoint.
\begin{lstlisting}
    POST <HOST_NAME>/query/timeseries
    JSON Body:
    {
        "geoJson": {
            "type": "Polygon",
            "coordinates": [
                [
                    [
                        <longitude>,
                        <latitude>
                    ],
                    ...
                ]
            ]
        },
        "parameter": "<PARAMETER_ID>"
    }
\end{lstlisting}

\paragraph{Query All Timeseries} -- Send a request with empty JSON body will respond with all the timeseries.
\begin{lstlisting}
    POST <HOST_NAME>/query/timeseries
    JSON Body:
    {}
\end{lstlisting}

\dbc{Need a chapter summary Sec.}
